Analysis of Bauer et al. (2023) bioRxiv: <br> Favourable Conservation Status (FCS)

### 2023-03-01 R/model_check_fcs.Rmd
Update model_check_fcs.Rmd

Markov Chain Monte Carlo (MCMC) method used with No-U-Turn Sampler
(NUTS) sampler which is an extension of Hamiltonian Monte Carlo (HMC).

@@

```{r possible-priors}
brms::get_prior(n ~ target_type + exposition + sand_ratio + survey_year_fct +
            seed_density + substrate_depth +
            (1 | site / plot) + (1 | botanist_year),
          data = sites)
data <- data.frame(x = c(-5, 5))
ggplot(data, aes(x = x)) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 2)) +
  expand_limits(y = 0) +
  ggtitle("Normal distribution for Intercept")
ggplot(data, aes(x = x)) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 0.3, sd = 2)) +
  expand_limits(y = 0) +
  ggtitle("Normal distribution for treatments")
ggplot(data, aes(x = x)) +
  stat_function(fun = dcauchy, n = 101, args = list(location = 0, scale = 1)) +
  expand_limits(y = 0) +
  ggtitle("Cauchy distribution")
ggplot(data, aes(x = x)) +
  stat_function(fun = dstudent_t, args = list(df = 3, mu = 0, sigma = 2.5)) +
  expand_limits(y = 0) +
  ggtitle(expression(Student ~ italic(t) * "-distribution"))
```

@@

```{r prior-summary}
brms::prior_summary(m_1, all = FALSE)
```

@@

```{r prior-predictive-check}
bayesplot::ppc_stat(y, yrep_prior[1:500, ], binwidth = 0.5) +
  coord_cartesian(xlim = c(-4, 4))
ppc_stat_grouped(
  y, yrep_prior[1:500, ], group = sites$site, binwidth = 0.5
  ) +
  coord_cartesian(xlim = c(-4, 4))
ppc_stat_grouped(
  y, yrep_prior[1:500, ], group = sites$exposition, binwidth = 0.5
  ) +
  coord_cartesian(xlim = c(-4, 4))
ppc_stat_grouped(
  y, yrep_prior[1:500, ], group = sites$survey_year_fct, binwidth = 0.5
  ) +
  coord_cartesian(xlim = c(-4, 4))
ppc_stat_grouped(
  y, yrep_prior[1:500, ], group = sites$target_type, binwidth = 0.5
  ) +
  coord_cartesian(xlim = c(-4, 4))
ppc_stat_grouped(
  y, yrep_prior[1:500, ], group = sites$seed_density, binwidth = 0.5
  ) +
  coord_cartesian(xlim = c(-4, 4))
ppc_stat_grouped(
  y, yrep_prior[1:500, ], group = sites$sand_ratio, binwidth = 0.5
  ) +
  coord_cartesian(xlim = c(-4, 4))
ppc_stat_grouped(
  y, yrep_prior[1:500, ], group = sites$substrate_depth, binwidth = 0.5
  ) +
  coord_cartesian(xlim = c(-4, 4))
```

@@

```{r mcmc-pairs, collapse = TRUE, fig.height = 6.5}
bayesplot::mcmc_pairs(m_1, off_diag_args = list(size = 1.2),
           pars = c(
             "b_sand_ratio25", "b_sand_ratio50", "b_substrate_depth15",
             "b_target_typedry_grassland", "b_seed_density8",
             "b_expositionsouth", "sigma"
           ))
bayesplot::mcmc_pairs(m_2, off_diag_args = list(size = 1.2),
           pars = c(
             "b_sand_ratio25", "b_sand_ratio50", "b_substrate_depth15",
             "b_target_typedry_grassland", "b_seed_density8",
             "b_expositionsouth", "sigma"
           ))
```

@@

```{r mcmc-parcoord, collapse = TRUE}
bayesplot::mcmc_parcoord(posterior1, np = hmc_diagnostics1) +
  theme(axis.text.x = element_text(angle = 45))
bayesplot::mcmc_parcoord(posterior2, np = hmc_diagnostics2) +
  theme(axis.text.x = element_text(angle = 45))
```

@@

```{r autocorrelation}
bayesplot::mcmc_acf(posterior1, lags = 10)
bayesplot::mcmc_acf(posterior2, lags = 10)
```

@@

```{r kernel-density}
p1 <- bayesplot::ppc_dens_overlay(y, yrep1[1:50, ])
p2 <- bayesplot::ppc_dens_overlay(y, yrep2[1:50, ])
p1 / p2
ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$site)
ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$site)
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$exposition)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$exposition)
p1 / p2
ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$survey_year_fct)
ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$survey_year_fct)
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$target_type)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$target_type)
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$seed_density)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$seed_density)
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$sand_ratio)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$sand_ratio)
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$substrate_depth)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$substrate_depth)
p1 / p2
```

@@

```{r histogram}
p1 <- bayesplot::ppc_stat(y, yrep1, binwidth = 0.001)
p2 <- bayesplot::ppc_stat(y, yrep2, binwidth = 0.001)
p1 / p2
ppc_stat_grouped(y, yrep1, group = sites$site, binwidth = 0.001)
ppc_stat_grouped(y, yrep2, group = sites$site, binwidth = 0.001)
p1 <- ppc_stat_grouped(y, yrep1, group = sites$exposition, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$exposition, binwidth = 0.001)
p1 / p2
ppc_stat_grouped(y, yrep1, group = sites$survey_year_fct, binwidth = 0.001)
ppc_stat_grouped(y, yrep2, group = sites$survey_year_fct, binwidth = 0.001)
p1 <- ppc_stat_grouped(y, yrep1, group = sites$target_type, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$target_type, binwidth = 0.001)
p1 / p2
p1 <- ppc_stat_grouped(y, yrep1, group = sites$seed_density, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$seed_density, binwidth = 0.001)
p1 / p2
p1 <- ppc_stat_grouped(y, yrep1, group = sites$sand_ratio, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$sand_ratio, binwidth = 0.001)
p1 / p2
p1 <- ppc_stat_grouped(y, yrep1, group = sites$substrate_depth, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$substrate_depth, binwidth = 0.001)
p1 / p2
```

@@

```{r loo}
loo1 <- m_1 %>% brms::loo(save_psis = TRUE, moment_match = FALSE)
loo2 <- m_2 %>% brms::loo(save_psis = TRUE, moment_match = FALSE)
loo1
loo2
plot(loo1)
plot(loo2)
```

@@

```{r loo-pit}
p1 <- bayesplot::ppc_loo_pit_overlay(y, yrep1, lw = weights(loo1$psis_object))
p2 <- bayesplot::ppc_loo_pit_overlay(y, yrep2, lw = weights(loo2$psis_object))
p1 / p2

```

@@

```{r dharma, collapse = TRUE}
DHARMa.helpers::dh_check_brms(m_1, integer = TRUE)
DHARMa.helpers::dh_check_brms(m_2, integer = TRUE)
```

@@

```{r conditional-r2, collapse = TRUE}
brms::bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site / plot))
brms::bayes_R2(m_2, probs = c(0.05, 0.5, 0.95),
         re_formula =  ~ (1 | site / plot))
```

@@

```{r marginal-r2, collapse = TRUE}
brms::bayes_R2(m_1, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
brms::bayes_R2(m_2, probs = c(0.05, 0.5, 0.95),
         re_formula = 1 ~ 1)
```

@@

```{r effect-sizes, message = FALSE}
posterior1 %>%
  posterior::summarize_draws() %>%
  knitr::kable()
emm <- emmeans(
  m_1, revpairwise ~ target_type + sand_ratio | exposition | survey_year_fct,
  type = "response"
  )
emm$emmeans
```

>>>>

Markov Chain Monte Carlo (MCMC) method used with No-U-Turn Sampler
(NUTS) which is an extension of Hamiltonian Monte Carlo (HMC).

@@

```{r possible-priors}
brms::get_prior(
  n ~ target_type + sand_ratio + seed_density + substrate_depth +
    exposition + survey_year_fct + botanist_year + (1 | site / plot),
  data = sites
  ) %>%
  select(prior, class, coef, group, source)
data <- data.frame(x = c(-5, 5))
ggplot(data, aes(x = x)) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 2)) +
  expand_limits(y = 0) +
  ggtitle("Normal distribution for Intercept")
ggplot(data, aes(x = x)) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 0.3, sd = 2)) +
  expand_limits(y = 0) +
  ggtitle("Normal distribution for treatments")
ggplot(data, aes(x = x)) +
  stat_function(fun = dcauchy, n = 101, args = list(location = 0, scale = 1)) +
  expand_limits(y = 0) +
  ggtitle("Cauchy distribution")
ggplot(data, aes(x = x)) +
  stat_function(fun = dstudent_t, args = list(df = 3, mu = 0, sigma = 2.5)) +
  expand_limits(y = 0) +
  ggtitle(expression(Student ~ italic(t) * "-distribution"))
```

@@

```{r prior-summary}
m_1 %>%
  brms::prior_summary(all = FALSE) %>%
  select(prior, class, coef, group, source)
```

@@

```{r prior-predictive-check, fig.height = 4}
bayesplot::ppc_stat(y, yrep_prior[1:500, ], binwidth = 0.5) +
  coord_cartesian(xlim = c(-4, 4))
ppc_stat_grouped(
  y, yrep_prior[1:500, ], group = sites$site, binwidth = 0.5
  ) +
  coord_cartesian(xlim = c(-4, 4))
ppc_stat_grouped(
  y, yrep_prior[1:500, ], group = sites$exposition, binwidth = 0.5
  ) +
  coord_cartesian(xlim = c(-4, 4))
ppc_stat_grouped(
  y, yrep_prior[1:500, ], group = sites$survey_year_fct, binwidth = 0.5
  ) +
  coord_cartesian(xlim = c(-4, 4))
ppc_stat_grouped(
  y, yrep_prior[1:500, ], group = sites$target_type, binwidth = 0.5
  ) +
  coord_cartesian(xlim = c(-4, 4))
ppc_stat_grouped(
  y, yrep_prior[1:500, ], group = sites$seed_density, binwidth = 0.5
  ) +
  coord_cartesian(xlim = c(-4, 4))
ppc_stat_grouped(
  y, yrep_prior[1:500, ], group = sites$sand_ratio, binwidth = 0.5
  ) +
  coord_cartesian(xlim = c(-4, 4))
ppc_stat_grouped(
  y, yrep_prior[1:500, ], group = sites$substrate_depth, binwidth = 0.5
  ) +
  coord_cartesian(xlim = c(-4, 4))
```

@@

```{r mcmc-pairs, collapse = TRUE, fig.height = 6.5}
m_1 %>% bayesplot::mcmc_pairs(
  off_diag_args = list(size = 1.2),
  pars = c(
    "b_sand_ratio25", "b_sand_ratio50", "b_substrate_depth15",
    "b_target_typedry_grassland", "b_seed_density8",
    "b_expositionsouth", "sigma"
           )
)
m_2 %>% bayesplot::mcmc_pairs(
  off_diag_args = list(size = 1.2),
  pars = c(
    "b_sand_ratio25", "b_sand_ratio50", "b_substrate_depth15",
    "b_target_typedry_grassland", "b_seed_density8",
    "b_expositionsouth", "sigma"
           )
)
```

@@

```{r mcmc-parcoord, collapse = TRUE}
posterior1 %>% bayesplot::mcmc_parcoord(np = hmc_diagnostics1) +
  theme(axis.text.x = element_text(angle = 45))
posterior2 %>% bayesplot::mcmc_parcoord(np = hmc_diagnostics2) +
  theme(axis.text.x = element_text(angle = 45))
```

@@

```{r autocorrelation}
posterior1 %>% bayesplot::mcmc_acf(lags = 10)
posterior2 %>% bayesplot::mcmc_acf(lags = 10)
```

@@

```{r kernel-density}
p1 <- bayesplot::ppc_dens_overlay(y, yrep1[1:50, ])
p2 <- bayesplot::ppc_dens_overlay(y, yrep2[1:50, ])
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$site)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$site)
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$exposition)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$exposition)
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$survey_year_fct)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$survey_year_fct)
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$target_type)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$target_type)
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$seed_density)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$seed_density)
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$sand_ratio)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$sand_ratio)
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$substrate_depth)
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$substrate_depth)
p1 / p2
```

@@

```{r histogram}
p1 <- bayesplot::ppc_stat(y, yrep1, binwidth = 0.001)
p2 <- bayesplot::ppc_stat(y, yrep2, binwidth = 0.001)
p1 / p2
p1 <- ppc_stat_grouped(y, yrep1, group = sites$site, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$site, binwidth = 0.001)
p1 / p2
p1 <- ppc_stat_grouped(y, yrep1, group = sites$exposition, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$exposition, binwidth = 0.001)
p1 / p2
p1 <- ppc_stat_grouped(y, yrep1, group = sites$survey_year_fct, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$survey_year_fct, binwidth = 0.001)
p1 / p2
p1 <- ppc_stat_grouped(y, yrep1, group = sites$target_type, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$target_type, binwidth = 0.001)
p1 / p2
p1 <- ppc_stat_grouped(y, yrep1, group = sites$seed_density, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$seed_density, binwidth = 0.001)
p1 / p2
p1 <- ppc_stat_grouped(y, yrep1, group = sites$sand_ratio, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$sand_ratio, binwidth = 0.001)
p1 / p2
p1 <- ppc_stat_grouped(y, yrep1, group = sites$substrate_depth, binwidth = 0.001)
p2 <- ppc_stat_grouped(y, yrep2, group = sites$substrate_depth, binwidth = 0.001)
p1 / p2
```

@@

```{r loo}
loo1 <- m_1 %>% brms::loo(save_psis = TRUE, moment_match = TRUE)
loo2 <- m_2 %>% brms::loo(save_psis = TRUE, moment_match = TRUE)
loo1
loo2
plot(loo1)
plot(loo2)
```

@@

```{r loo-pit}
p1 <- bayesplot::ppc_loo_pit_overlay(y, yrep1, lw = weights(loo1$psis_object))
p2 <- bayesplot::ppc_loo_pit_overlay(y, yrep2, lw = weights(loo2$psis_object))
p1 / p2
```

@@

```{r dharma, collapse = TRUE}
m_1 %>% DHARMa.helpers::dh_check_brms(integer = TRUE)
m_2 %>% DHARMa.helpers::dh_check_brms(integer = TRUE)
```

@@

```{r conditional-r2, collapse = TRUE}
m_1 %>% brms::bayes_R2(
  probs = c(0.05, 0.5, 0.95), re_formula =  ~ (1 | site / plot)
  )
m_2 %>% brms::bayes_R2(
  probs = c(0.05, 0.5, 0.95), re_formula =  ~ (1 | site / plot)
  )
```

@@

```{r marginal-r2, collapse = TRUE}
m_1 %>% brms::bayes_R2(
  probs = c(0.05, 0.5, 0.95), re_formula = 1 ~ 1
  )
m_2 %>% brms::bayes_R2(
  probs = c(0.05, 0.5, 0.95), re_formula = 1 ~ 1
  )
```

@@

```{r effect-sizes, message = FALSE}
posterior1 %>%
  posterior::summarize_draws() %>%
  knitr::kable() #%>%
  #kableExtra::kable_styling(font_size = 10)
emm <- m_1 %>% emmeans(
  revpairwise ~ target_type + sand_ratio | exposition | survey_year_fct,
  type = "response"
  )
emm$emmeans
```
